---
title: "Men's Cross Country Ski 2010-2022 Olympics"
output: html_document
date: "2023-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggiraphExtra)
```

```{r}
last_four_cc <- read_csv("data/mens_cc_ski_2010-22.csv")
```

```{r}
ggplot(data = last_four_cc,
       aes(x = factor(Year), y = Time_min, fill = factor(Year)))+
  geom_boxplot()
```

```{r}
year_mod = aov(Time_min ~ factor(Year), data = last_four_cc)
summary(year_mod)

year_mod$coefficients
model.tables(year_mod, type = "means")
model.tables(year_mod, type = "effects")
```

```{r}
TukeyHSD(year_mod)

ggHSD( TukeyHSD(year_mod) , no = 1)
```
# do two sample t tests for most significant
# find other ways to compare


```{r}
last_four_cc <- last_four_cc%>%
  mutate(scandinavian = ifelse(Country == "SWE"|Country == "DEN"|Country == "NOR"|Country == "FIN"| Country == "ISL"|Country == "ISL", 1, 0))
```

# Density plot
```{r}
ggplot(data = last_four_cc, 
       aes(x = Rank, fill = factor(scandinavian)))+
  geom_density(alpha = .6)

# 1 = Scandinavian
# 0 = not Scandinavian
```
# Boxplot

```{r}
ggplot(data = last_four_cc, 
       aes(x = scandinavian, y = Rank, fill = factor(scandinavian)))+
  geom_boxplot()
```

# Unpaired two sample t-test for Scandinavians
```{r}
last_four_cc %>%
with(t.test(Rank ~ factor(scandinavian)))

# highly significant difference in means
```

# Looking at comparisons of means by year 

```{r}
last_four_cc%>%
  group_by(Year, scandinavian)%>%
  summarise("Mean Rank" = mean(Rank)) 
# 2010 difference in means: 20.38 
# 2014 difference in means: 28.60
# 2018 difference in means: 38.10
# 2022 differnce in means: 36.54
```
# Looking at numbers by year 
```{r}
last_four_cc%>%
  group_by(Year, scandinavian)%>%
  summarise("Total Athletes" = n())
# 2010 prop Scandinavian: 13/82 = .16
# 2014 prop Scandinavian: 14/73 = .19
# 2018 prop Scandinavian: 14/102 = .14
# 2022 prop Scandinavian: 13/82 = .16

# Change is not too significant 
```
# possibly find way to inccorporate indicator variable
# Though elements of the courses vary year to year due to factors like elevation change, elevation of the city, conditions of the snow, etc., time is still a decent predictor for rank
```{r}
cor(last_four_cc$Time_min, last_four_cc$Rank)

mod1 <- lm(Rank ~ Time_min, data = last_four_cc)
summary(mod1)
```

# Adding the indicator variable of scandinavian
```{r}
mod2 <- lm(Rank ~ scandinavian + Time_min, data = last_four_cc)
summary(mod2)
```



