---
title: "Mens 15k 2022"
output: html_document
date: "2023-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
cross_country <- read_csv("data/mens_15k_beijing_2022.csv")
```

# scatterplot of rank vs result times in minutes
```{r}
ggplot(data = cross_country, 
       aes(x = Rank, y = Time_min))+
  geom_point()
```
# summary statistics of result times in minutes
```{r}
summary(cross_country$Time_min)
```
# outlier visualizations
```{r}
plot(cross_country$Time_min)
hist(cross_country$Time_min)
boxplot(cross_country$Time_min)

```
# calculating thresholds with mean and std
```{r}
mean = mean(cross_country$Time_min)
std = sd(cross_country$Time_min)

# get threshold values for outliers
Tmin = mean-(3*std)
Tmax = mean+(3*std)

# find outlier
cross_country$Time_min[which(cross_country$Time_min < Tmin | cross_country$Time_min > Tmax)]

# [1] 55.69833 --- corresponds to last place finisher Carlos Andrés Quintana
```

# calculating mild outlier thresholds with IQR and quartiles
```{r}
iqr = IQR(cross_country$Time_min)


# get threshold values for mild outliers
Tmin = 40.66-(1.5*iqr) 
Tmax = 44.35+(1.5*iqr) 

# find mild outliers
cross_country$Time_min[which(cross_country$Time_min < Tmin | cross_country$Time_min > Tmax)]

# several: 50.585, 50.610, 51.775, 52.748, 53.500, 55.698
```
# calculating extreme outlier thresholds with IQR and quartiles
```{r}
iqr = IQR(cross_country$Time_min)


# get threshold values for mild outliers
Tmin = 40.66-(3*iqr) 
Tmax = 44.35+(3*iqr) 

# find mild outliers
cross_country$Time_min[which(cross_country$Time_min < Tmin | cross_country$Time_min > Tmax)]

# [1] 55.69833 --- same runner identified again
```
# Grubb's Test
```{r}
grubbs.test(cross_country$Time_min)
```
# calculating z-score of Carlos Andrés Quintana
```{r}
# scale(cross_country$Time_min)

z_quintana = (55.69833-43.01) / 3.553
# 3.571
# Quite high when put in context.
# In contrast, the fastest runner who won gold had a z-score of -1.434
```


# Simple linear regression of Time vs. Rank
```{r}
rank_mod <- lm(Time_min ~ Rank, data = cross_country)
summary(rank_mod)
```
# Inference of model
On average, a 1 place increase in rank is associated with a 0.12 minute increase in result time for Men's Olympic 15k cross country skiers.

It is important to note that this is not a random sample but instead is taken directly from the 2022 summer Olympics in Beijing. This means it cannot be assumed that this model will be an accurate predictor for all Men's Olympic 15k cross country ski races, and likely is not, due to the fact that skiers tend to get faster over time throughout history. 


