---
title: "scandinavian module idea"
output: html_document
date: "2023-06-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# data
```{r}
last_four_cc <- read_csv("data/mens_cc_ski_2010-22.csv")
```

# creating 2 level variable for scandinavian vs. not scandinavian
```{r}
last_four_cc <- last_four_cc%>%
  mutate(scandinavian = ifelse(Country == "SWE"|Country == "DEN"|Country == "NOR"|Country == "FIN"| Country == "ISL", 1, 0))
```

# Density plot of times in minutes
```{r}
ggplot(data = last_four_cc, 
       aes(x = Time_min, fill = factor(scandinavian)))+
  geom_density(alpha = .6)+
  facet_wrap(vars(Year))

# 1 = Scandinavian
# 0 = not Scandinavian

# seems to be fairly notable difference
```

# two sample t-test and confidence interval
```{r}
last_four_cc %>%
with(t.test(Time_min ~ factor(scandinavian)))

# highly significant difference in means
```

# Time in minutes by if they are scandinavian or not

95 percent confidence interval:
-On average, and with 95% confidence, the difference in mean result times for Scandinavians and non-Scandinavians in between 1.98 3.81 minutes

-On average Scandinavian cross country skiers finish the 15km race 1.98 to 3.81 minutes faster than the rest of the field. 


# Significant, but does it make sense to interpret times when you take into account that the courses vary considerably from year to year?
# Snow quality, elevation change, altitude, etc.
```{r}
ggplot(data = last_four_cc,
       aes(x = factor(Year), y = Time_min, fill = factor(Year)))+
  geom_boxplot()

# no clear pattern, with medians changing by Olympic year
```

# Would it make more sense to look at rank results instead?
# comparing times by year to see variation

# Density plot of difference in ranks
```{r}
ggplot(data = last_four_cc, 
       aes(x = Rank, fill = factor(scandinavian)))+
  geom_density(alpha = .6)+
  facet_wrap(vars(Year))

# 1 = Scandinavian
# 0 = not Scandinavian

# seems to be a more clear difference
```

# Looking at comparisons of means by year 
```{r}
last_four_cc%>%
  group_by(Year, scandinavian)%>%
  summarise("Mean Rank" = mean(Rank)) 
# 2010 difference in means: 20.38 
# 2014 difference in means: 28.60
# 2018 difference in means: 38.10
# 2022 differnce in means: 36.54 

# Large, and has grown each even (with exception to 2022 but still large)
```
# two-sample t-test and confidence interval

```{r}
last_four_cc %>%
with(t.test(Rank ~ factor(scandinavian)))

# Highly significant again with clear implications
```
# On average, and with 95% confidence, the ranks of scandinavian 15km cross country athletes are between 22.33 and 37.88 places higher than the rest of the field. 

# Is using rank better for analysis? What do you lose and what do you gain by using rank instead of time?
# Gain more objective viewing of performance since course median times vary year to year due to course variation.
# Lose difference in scale of times because rank does not account for how different the times are from rank to rank(eg. 1-2 is same as 12-13)




# Now, would it make more sense to look at z-score results instead?

