---
title: "boston_zscore"
output: html_document
date: "2023-06-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Calculating z-scores for top Male and Female finishers to see which was more remarkable in comaprison to the field

```{r}
boston_Men<-boston%>%
  filter(gender == "M")

write_csv(x = boston_Men, file = "data/boston_men.csv")

boston_Women <- boston%>%
  filter(gender == "F")

write_csv(x = boston_Women, file = "data/boston_women.csv")
```
## Histograms of the results respectively for male and female

```{r}
hist(boston_Men$Result_sec)
hist(boston_Women$Result_sec)
```

# Finding mean and standard deviation for male runners and calculating z-score

```{r}
# find mean and standard deviation of MENS results in seconds
boston_Men%>%
  summarise(mean_result_sec = mean(Result_sec),
            sd_result_sec = sd(Result_sec))

# mean = 13377.54
#sd = 2737.843

# 1st place male runner
boston_Men%>%
  slice_min(Result_sec)

zscore <- (7677 - 13377.54) / 2737.843
zscore
#z-score of -2.082128
```

# Finding mean and standard deviation for female runners and calculating z-score
```{r}
boston_Women%>%
  summarise(mean_result_sec = mean(Result_sec),
            sd_result_sec = sd(Result_sec))

# mean = 14717.8
#sd = 2454.376

# 1st place female runner
boston_Women%>%
  slice_min(Result_sec)

zscore <- (8611 - 14717.8) / 2454.376
zscore
# z-score of -2.488127
```

# In conclusion, the top female finishers result was more remarkable for that day in relation to the female field
# than the top male finisher because the score was further from zero.