---
title: "Boston Marathon 2019"
author: "Jack Fay"
date: "2023-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Motivation

The Boston Marathon, a legendary event etched in the hearts of runners and 
spectators alike,represents the epitome of resilience, perseverance, and 
triumph. Year after year, participants from around the world gather, united by
a shared passion for testing their limits and embracing the relentless pursuit 
of greatness. In 2019, this storied race witnessed the rise of two extraordinary
champions who left an indelible mark on history. On the men's side, 
Lawrence Cherono, a powerhouse of determination, unleashed his formidable 
strength, pushing his body to its limits to claim victory with unparalleled 
resolve. Meanwhile, in the women's division, Worknesh Degefa demonstrated a 
relentless spirit, blazing through the course with unwavering determination, 
leaving her competitors in awe as she crossed the finish line triumphantly. 
Both Lawrence Cherono and Worknesh Degefa exemplify the very essence of the 
Boston Marathon, where dreams are transformed into reality through perseverance,
grit, and unwavering belief. Their triumphs serve as a powerful reminder that
with unwavering dedication and unyielding determination, we too can overcome any
obstacle that stands in our way, and cross our own personal finish lines, 
victorious in our pursuit of greatness.

## Data

The data set is from the 2019 Boston Marathon and contains observations for all finishers of the race.
The 'boston' data set contains 26651 total observations with each row representing a runner. 


Variables:
*Rank_Tot - overall ranking in the race
*Age - age of the runner 
*Country - nationality of the runner 
*Result_hr - finishing time of the runner in hours minutes seconds format
*Result_sec - finishing time of the runner in seconds 
*Country_code - country abbreviation 
*gender - gender of the runner (male or female)

```{r}
boston <- read_csv("~/Downloads/Dataset-Boston-2019.csv")

write_csv(x = boston, file = "data/Dataset-Boston-2019.csv")
```

```{r}
boston<-boston%>%
  mutate(gender = factor(Gender))%>%
  select(-Gender)%>%
  na.omit()

head(boston)
```

## Is there a statistical difference in means for race results for males and females?


## Side-by-side boxplot displaying gender vs result in seconds 
```{r}
ggplot(data = boston, 
        aes(x = gender, Result_sec, fill = gender))+
  geom_boxplot()
```

## Two Sample t-test to determine whether there is a difference in means
```{r}
boston %>%
with(t.test(Result_sec ~ gender))

# yes, the difference in means is statistically significant 
```









