---
title: "Cleaning"
output: html_document
date: "2023-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


# There is a significantly higher correlation between faceoff win percentage and overall win percentage for women than there is for men

# Shot percetage on goal also has a strong correlation with win percentage 

```{r}
draw <- read_csv("draw_pctg.csv")
wins <- read_csv("win_pctg.csv", col_select = 1:2)
turnovers <- read_csv("turnovers.csv", col_select = 1:2)
sog <- read_csv("sog.csv", col_select = 1:2)


w_lax <- left_join(draw, wins, by = "Team")
w_lax <- left_join(w_lax, turnovers, by = "Team")
w_lax <- left_join(w_lax, sog, by = "Team")


cor(w_lax$draw_pctg, w_lax$win_pctg)
cor(w_lax$sog_per_game, w_lax$win_pctg)
```

```{r}
mod <- lm(win_pctg ~ sog_per_game + draw_pctg, data = w_lax)
summary(mod)
```


# Idea: Have students compare the correlation of of faceoff_win_pctg to win_pctg for men and women
# Its higher for women being the faceoffs are more meaningful.

# Then, why could this be?
# look at histograms of turnovers per game for men and women 

```{r}
write_csv(w_lax, file = "womens_lax_2022-2023.csv")

w_lax <- read_csv("womens_lax_2022-2023.csv")


# Correlation is higher for women than men 
cor(w_lax$draw_pctg, w_lax$win_pctg)
cor(lax$faceoff_win_pct, lax$win_loss_pctg)
```
```{r}
# 
hist(lax$avg_turnovers)
hist(w_lax$turnovers_per_game)
```
```{r}
mean(lax$avg_turnovers)
mean(w_lax$turnovers_per_game)

# not crazy
```

