---
title: "Disc Golf League"
output: html_document
date: "2023-06-16"
---
# Website "StatMando" has wide variety of disc golf data 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(readr)
```

# Scoring data by hole from 2023 MPO Standard Elite Series & Majors for all players

Before the Pro Tour, there were limited opportunities for our sport’s top athletes to compete in dedicated events. These players have the greatest potential to pursue a career in professional touring, and as such, our premier events should prioritize them and provide them with greater visibility. The Pro Tour exclusively features two Open divisions: Men’s (MPO) and Women’s (FPO). This ensures that all of the added cash at the event goes directly to these players, providing them with an opportunity to tour full-time. Moreover, the Pro Tour collaborates with players to facilitate local and on-site clinics, generate media exposure, and supply comprehensive season-long statistics through our partnership with UDisc.
```{r}
url <- "https://statmando.com/stats/tour-holes-2023-mpo"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

#write_csv(x =  objs[[1]], file = "disc_golf_holes.csv")
#disc <- read_csv("disc_golf_holes.csv", skip = 2)
#write_csv(x =  disc, file = "disc_golf_holes.csv")


disc <- read_csv("disc_golf_holes.csv")
disc<-disc%>%
  na.omit()
```
uP = under Par (birdies or better).
oP = over Par (bogeys or worse). 
CiRDS = Circle in Regulation Distance per Shot. (distance required per shot to hit circle in regulation)

Expected counts for over-par, par, and under-par out of times played. Should be 1/3*times_played

# Hypothesis Testing: One categorical variable 



# Chi-square goodness of fit

```{r}
disc[1,]
```
# Prepare
# should null proportions be same as cumulative proportions for the data?
# or decided by general thought process .25, .5, .25
```{r}
disc%>%
  summarise(prop_uPar = sum(`# uPar`)/sum(`Times Played`),
            prop_Par = sum(`# Pars`)/sum(`Times Played`),
            prop_oPar = sum(`# oPar`)/sum(`Times Played`))
```

Hole #16 at the OTB Open
- Seeing if there is an equal amount of scores under-par, over-par, and of par

- score can be under-par/par/over-par (3 level categorical variable)

If these scores were decided by random chance, approximately how often would there be a par (or under/over)?
- 
Ho: hole scores decided by random chance (null value = 1/3)
Ha: results are not determined by random chance 


Ho: P-under = .25
    P-par   = .5
    P-over  = .25

# Check for independence 
All scores from different players 

# Calculate 
If the null hypothesis were true, what would we expect our counts to be?
Determine expected counts for distribution.

- E-under = 329*.25 = 82.25
- E-par = 329*.5 = 164.5
- E-over = 329*.25 = 82.25 


# Sample size condition 
All expected counts should be larger than 5.
- Clearly satisfied 


# Find X^2 value and p-value 

X^2 = (30 - 82.25)^2/82.25 + (155 - 164.5)^2/164.5 + (144 - 82.25)^2/82.25

X^2 = 33.19 + .54 + 46.36 = 80.09

Degrees of freedom = (# of levels in variable) - 1
df = 2

https://www.lock5stat.com/StatKey/theoretical_distribution/theoretical_distribution.html#chi

p-value = 0.000

```{r}
pchisq(q=80.09, df=2, lower.tail=FALSE)
```


# Conclude 

- We estimate the proportion of people who have scores under-par, of par,
 and over-par to be 30/329, 155/329, and 144/329 respectively. 
 There is very strong evidence that at least one of the proportions is different than the average, (X^2 = 87.37, p-value = .0000).

Which level was the biggest contributor to the chi-square statistic?
- over-par

Why do you think that is?

- Because it is seemingly a difficult hole to score under-par on. It is also more difficult than average to make a par.

## It is evident that scores of par are common but following the assumed proportions, the hole is more challenging than average. It causes players to be over par more frequently, and does not allow for normal frequencies of under par scores.


# This can be done with lots of different holes so they can be compared 

------------------------------
# Now 

# Chi-square test of association
Could be done by comparing:
-pars 3,4,5
-Events 
-by year for same event
-etc.
```{r}
disc%>%
  filter(Event == "OTB Open")%>%
  summarise(uPar = sum(`# uPar`),
            Par = sum(`# Pars` ),
            oPar = sum(`# oPar`),
            Times_played = sum(`Times Played`))

disc%>%
  filter(Event == "Las Vegas Challenge")%>%
  summarise(uPar = sum(`# uPar`),
            Par = sum(`# Pars` ),
            oPar = sum(`# oPar`),
            Times_played = sum(`Times Played`))
```

# Variables of interest 
Event/Course(one year so same course) - 2 levels
Score on the holes - 3 levels

Ho: no association between event and score on holes 
Ha: association

expected counts:













