---
title: "Disc Golf League"
output: html_document
date: "2023-06-16"
---
# Website "StatMando" has wide variety of disc golf data 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```

# Scoring data from 2023 MPO Standard Elite Series & Majors for all players
```{r}
url <- "https://statmando.com/stats/tour-holes-2023-mpo"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

#write_csv(x =  objs[[1]], file = "disc_golf_holes.csv")
#disc <- read_csv("disc_golf_holes.csv", skip = 2)
#write_csv(x =  disc, file = "disc_golf_holes.csv")


disc <- read_csv("disc_golf_holes.csv")
disc<-disc%>%
  na.omit()
```
uP = under Par (birdies or better).
oP = over Par (bogeys or worse). 
CiRDS = Circle in Regulation Distance per Shot. (distance required per shot to hit circle in regulation)

Topics to explore: 
MLR 
Nested F-tests 
Expected counts for over-par, par, and under-par out of times played. Should be 1/3*times_played
# possibly make an indicator variable for Length if it exceed a certain amount. longIND

```{r}
cor(disc[, 3:15])

```

```{r}
cor(disc$Length, disc$Avg)

mod1 <- lm(Avg ~ Length, data = disc)
summary(mod1)

mod2 <- lm(Avg ~ Length, data = disc)
summary(mod2)
```

