---
title: "Disc Golf League"
output: html_document
date: "2023-06-16"
---
# Website "StatMando" has wide variety of disc golf data 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```

# Scoring data from 2023 MPO Standard Elite Series & Majors for all players
```{r}
url <- "https://statmando.com/stats/tour-holes-2023-mpo"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

#write_csv(x =  objs[[1]], file = "disc_golf_holes.csv")
#disc <- read_csv("disc_golf_holes.csv", skip = 2)
#write_csv(x =  disc, file = "disc_golf_holes.csv")


disc <- read_csv("disc_golf_holes.csv")
disc<-disc%>%
  na.omit()
```
uP = under Par (birdies or better).
oP = over Par (bogeys or worse). 
CiRDS = Circle in Regulation Distance per Shot. (distance required per shot to hit circle in regulation)

Topics to explore: 
MLR 
Nested F-tests 
Expected counts for over-par, par, and under-par out of times played. Should be 1/3*times_played
# possibly make an indicator variable for Length if it exceed a certain amount. longIND

```{r}
cor(disc[, 3:15])

```

```{r}
cor(disc$Length, disc$Avg)

mod1 <- lm(Avg ~ Length, data = disc)
summary(mod1)

mod2 <- lm(Avg ~ Length, data = disc)
summary(mod2)
```
# Hypothesis Testing: One categorical variable 
```{r}
disc[1,]
```
# Prepare
Hole #16 at the OTB Open
- score can be under-par/par/over-par (3 level categorical variable)

If these scores were decided by random chance, approximately how often would there be a par (or under/over)?
- with random chance each would occur roughly 1/3 of the time 
(sampling variability makes it non-exact to 1/3)

Ho: hole scores decided by random chance (null value = 1/3)
Ha: results are not determined by random chance 


Ho: P-under = 1/3
    P-par   = 1/3
    P-over  = 1/3

# Check for independence 
All scores from different players 

# Calculate 
If the null hypothesis were true, what would we expect our counts to be?
Determine expected counts for dstribution.

- E-under = 329*1/3 = 109.67
- E-par = 329*1/3 = 109.67
- E-over = 329*1/3 = 109.67


# Sample size condition 
All expected counts should be larger than 5.
- Clearly satisfied because 109.67 > 5


# Find X^2 value and p-value 

X^2 = (30 - 109.67)^2/109.67 + (155 - 109.67)^2/109.67 + (144 - 109.67)^2/109.67

X^2 = 57.88 + 18.74 + 10.75 = 87.37

Degrees of freedom = (# of levels in variable) - 1
df = 2

https://www.lock5stat.com/StatKey/theoretical_distribution/theoretical_distribution.html#chi

p-value = 0.000

```{r}
pchisq(q=87.37, df=2, lower.tail=FALSE)
```


# Conclude 

- We estimate the proportion of people who have scores under-par, of par,
 and over-par to be 30/329, 155/329, and 144/329 respectively. 
 There is very strong evidence that at least one of the proportions is different than 
 1/3 (X^2 = 87.37, p-value = .0000)

Which level was the biggest contributor to the chi-square statistic?
- Under-par

Why do you think that is?

- Because it is seemingly a difficult hole to score under par on. Although, 
more people make par than over-par

Perhaps it is a par 3 which makes it tricky 
to make it on the second shot?









