---
title: "Disc Golf"
output: html_document
date: "2023-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
disc<- read_csv("disc_golf_holes.csv")

disc<-disc|>
  na.omit()
```


Topics to explore: 
MLR 
Nested F-tests 




# possibly make an indicator variable for Length if it exceed a certain amount. longIND

```{r}
cor(disc[, 3:15])

```

```{r}
cor(disc$Length, disc$Avg)

```

```{r}
ggplot(data = disc,
       aes(x = factor(Event), y = Avg, color = Event))+
  geom_point()
```
# using average to par to predict Avg score for a hole

```{r}
avg_mod <- lm(Avg ~ `Avg to Par`, data = disc)
summary(avg_mod)

```
# Creating an indicator variable for holes with lengths below and above the mean length
```{r}
disc%>%
  summarise(mean_length = mean(Length))

disc<-disc%>%
  mutate(LengthIND = ifelse(Length > 527,1, 0))
```

# Indicator variable model for intercept
```{r}
avg_mod2 <- lm(Avg ~ LengthIND + `Avg to Par`, data = disc)
summary(avg_mod2)

# model improves greatly
```
# Indicator variable model for intercept and slope

```{r}
avg_mod3 <- lm(Avg ~ LengthIND + `Avg to Par`*LengthIND, data = disc)
summary(avg_mod3)

# Doesn't seem to improve the model
```
# Nested F-test to determine which model is better
```{r}
anova(avg_mod2, avg_mod3)

# avg_mod2 is clearly better 
```
# Model with Length as a predictor for Avg score

```{r}
avg_mod4 <- lm(Avg ~ Length, data = disc)
summary(avg_mod4)

# best model
```

