---
title: "Disc Golf League"
output: html_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```
# Tournament: DGPT Elite+ 2023 Portland Open presented by Latitude 64
```{r}
url <- "https://www.pdga.com/tour/event/69022"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

mens<- objs[[2]]
womens<- objs[[3]]

portland_open <- 
```

# 2019 MPO World Rankings - Year End
```{r}
url <- "https://www.pdga.com/players/rankings/world?year=2019&division=MPO"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

mens_world_rankings <- objs[[1]]

toDelete <- seq(0, length(mens_world_rankings), 2)
mens_world_rankings <- mens_world_rankings[-toDelete ,]


```

WRPR = World Rankings Player Rating
WRPR is the average of round ratings from B-tier and higher events in the 12 months prior to the World Ranking update.
RR = Ranking based on WRPR
Worlds = Ranking based on 2019 Pro Worlds
USDGC/USWDGC = Ranking based on 2019 USDGC / 2019 USWDGC
Euro = Ranking basked on 2019 European Open
NET = Average of a player's best finishes in National Tour and EuroTour events during the past 12 months
Best four finishes for MPO, best three finishes for FPO
DGPT = Disc Golf Pro Tour - Average of 4 best finishes during past 12 months

```{r}
rank_mod <- lm(Rank ~ NET)
```

