---
title: "Inference"
output: html_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggiraphExtra)
library(lmtest)
```

```{r}
boston <- read_csv("data/boston_marathon_2023.csv")
boston<-boston%>%
  mutate(across(c(8,9,10), ~ hms(.)))
  

```
# Confidence Intervals and Prediction Intervals 
# issues with inference due to non-constant variance 
```{r}
time_mod <- lm(finish_net ~ half_time, data = boston)
anova(rating_mod)
# compute R-squared value by hand 
# SSM/SST = 549242/640259 = 0.8578435 = R-squared
```
# plots for model assumptions
```{r}
plot(rating_mod)
ggPredict(rating_mod)
bptest(rating_mod)
# normality - Q-Q plot of residuals --- looks good
# linearity - residuals vs. fitted flat or not---- yes
# constancy of variance (homoscedasticity)(equal variance) - scatterplot and residuals vs. fitted --- does not look good
```

# Predict the driver rating of drivers with a quality pass percentage of 42 
```{r}
summary(rating_mod)
# Rating = 33.27907 + 1.01381*42 = 75.85909
# The model predicts drivers with a 42.0 percent quality pass percentage to have a driver rating of 75.86
```
# confidence interval for model predictor of Percent Quality Passes
```{r}
confint(rating_mod, level = .95)
# On average, and with 95% confidence, a one percent increase in percent quality passes is associated with an increase in driver rating between 0.99 and 1.04 points. 
```

# Compute a 95% confidence interval for the mean rating of all drivers with a percent of quality passes of 42
```{r}
newx = data.frame(PercentQualityPasses = 42.0)
predict(rating_mod, newdata = newx, interval = "conf", level = 0.90)
# On average, and with 95% confidence, a 42% quality pass percentage will result in a driver rating between 75.33 and 76.39.
```

# Compute a 95% prediction interval for the mean rating of all drivers with a percent of quality passes of 42
```{r}
newx = data.frame(PercentQualityPasses = 42.0)
predict(rating_mod, newdata = newx, interval = "pred", level = 0.90)
# The model predicts that 95% of drivers with a 42% quality pass percentage will result in a driver rating between 62.03 and 89.69.
```