---
title: "tennis"
output: html_document
date: "2023-06-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```
# Sumo wrestling terms:
https://en.wikipedia.org/wiki/Glossary_of_sumo_terms

-Data is from Sumo Reference and has a Banzuke for each event.



-Banzuke | a document listing the rankings of professional sumo wrestlers published before each official tournament (honbasho).

-Torikumi |  A bout during a tournament

-Yusho | A tournament championship in any division, awarded to the wrestler who wins the most bouts.

-Sansho | 'Three prizes'. Special prizes awarded to makuuchi wrestlers for exceptional performance.

-Rikishi | The ring name of a the sumo wrestler

-Shikona is name

-Official professional sumo tournament, called a honbasho, held six times a year since 1958, where the results affect the wrestlers' rankings.



# Natsu 2023 
# Makuuchi Banzuke 

#  Website has section where data is accessible for each major tournament of every year for all divisions:
#  data includes, wins, losses, weight, height, birth date


# Cleaning one 
```{r}
url <- "https://sumodb.sumogames.de/Banzuke.aspx?b=202305&heya=-1&shusshin=-1&bd=on&w=on&hl=on#legend"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()


write_csv(x = objs[[6]], file = "data/natsu_2023.csv")
natsu_east <- read_csv("data/natsu_2023.csv", col_select = 1:5)
natsu_west <- read_csv("data/natsu_2023.csv", col_select = 5:9)

natsu_west<-natsu_west%>%
  relocate( 5, .before = 1)%>%
  relocate(2, .after = 5)%>%
  rename(East = West)

natsu <- bind_rows(natsu_east, natsu_west)

natsu<- natsu %>%
  na.omit()%>%
  rename(name = East)%>%
  separate(col = Result, sep = "-", into = c("wins", "losses"))%>%
  mutate(wins = parse_number(wins),
      losses = parse_number(losses))%>%
  separate(col = `Height/Weight`, sep = "cm", into = c("height_cm", "weight_kg"))%>%
  mutate(height_cm = parse_number(height_cm),
         weight_kg = parse_number(weight_kg))

write_csv(x = natsu, file = "data/natsu_2023.csv")
```


# Next step:
# Write function that scrapes all the data: will need to change the year 
# Link has number that goes year then goes up by 2 for each tournament
# eg link: "blahblah202201" next: "blahblah202203"

issue is there are 6 a year 

so link goes from 01, 03, 05, ... to 11, 

