---
title: "tennis"
output: html_document
date: "2023-06-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Sumo wrestling terms:
https://en.wikipedia.org/wiki/Glossary_of_sumo_terms

-Data is from Sumo Reference and has a Banzuke for each event.

-A Banzuke is a document listing the rankings of professional sumo wrestlers published before each official tournament (honbasho).

-Official professional sumo tournament, called a honbasho, held six times a year since 1958, where the results affect the wrestlers' rankings.

-This data is from the 2023 Haru
```{r}
url <- "https://sumodb.sumogames.de/Banzuke.aspx?b=202303&heya=-1&shusshin=-1&h=on&sh=on&bd=on&hd=on&su=on&w=on&hr=on&ch=on&cs=on&cr=on&spr=on&sps=on&snr=on&sns=on&hl=on&simple=on"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

sumo <- objs[[1]]

sumo<-sumo%>%
  as.data.frame()%>%
  select(1:37)%>%
  slice(17:430)

#write_csv(x = sumo, file = "haru_2023.csv")

#haru <- read_csv("haru_2023.csv", skip = 1)

#write_csv(x = haru, file = "haru_2023.csv")

haru <- read_csv("haru_2023.csv")

haru<-haru%>%
  slice(-305)

haru<-haru%>%
  separate(col = `Height/Weight...9`, into = c("height_cm", "weight_kg"), sep = "cm")%>%
  separate(col = weight_kg, into = "weight_kg", sep = 4)%>%
  mutate(height_cm = parse_number(height_cm),
         weight_kg = parse_number(weight_kg))
```

W	L	East	Heya	Shusshin	Birth Date	Hatsu	University	Height/Weight	High	Career	Career Record	Prev	Prev W	Prev L	Next	Next W	Next L	Rank	West	Heya	Shusshin	Birth Date	Hatsu	University	Height/Weight	High	Career	Career Record	Prev	Prev W	Prev L	W	L	Next	Next W	Next L