---
title: "winter olympic prep"
---

Load a few packages
```{r}
library(dplyr)
library(stringr)
library(rvest)
```

Olympic Database: https://www.olympiandatabase.com/

Click on the "Results" tab and load
1) The Olympics of choice
2) The Sport
3) The Event

e.g., 
Beijing 2002 > Alpine Skiing > Women's Giant Salom

You should see a table with the event's results


Next, scrape the page
```{r}
url <- "https://www.olympiandatabase.com/index.php?id=109755&L=1"


all_tables <- 
  url |> 
    read_html() |>
    html_elements("table")

# inspect object -- notice that entry 4 is the table we want

```

Now it becomes tricky and annoying
Use this to get the winner
```{r}
all_text <- html_text2(all_tables[[4]])

text_version <- all_text |> str_split_1(pattern = '\n')

winner <- text_version[which(str_detect( text_version, " ") )[1]] |> str_split_1("\t") |> matrix(nrow = 1) |> as.data.frame() |>
  select(1,2,4,6) |>
  rename(place = V1,
         name = V2,
         country = V4,
         time = V6
         )
```


Use this to get all rows except the winner
```{r}
no_winner_row <-
  all_tables |>
      html_table()

# inspect object -- notice that entry 4 is the table we want

# do some cleaning
no_winner_row <- 
  no_winner_row[[4]] |> 
  filter(X2 != "") |> 
  slice(-1) |>
  select(1,2,6,8) |>
  rename(place = X1,
         name = X2,
         country = X6,
         time = X8
         )

```

Finally, put things back together

```{r}
alpine_skiing_w_giant_salom <- bind_rows(winner, no_winner_row)
```

Can finish cleaning from here.

Possible questions

