---
title: "NRL Teams"
output: html_document
date: "2023-07-07"
possible topics: SLR, estimation and prediction, correlation and visualizations
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
url <- "https://www.foxsports.com.au/nrl/nrl-premiership/stats/teams"
objs <- read_html(url)%>%
html_nodes("table")%>%
html_table()

write_csv(x = objs[[8]], file = "rugby_teams.csv")
nrl <- read_csv("rugby_teams.csv")
```

# 2023 STANDINGS THUS FAR 

```{r}
nrl <- nrl%>%
  mutate(`FG%` = parse_number(`FG%`))
nrl1<-
  nrl%>%
  select(-1)
cor(nrl1)

# PTS correlations: T, RM, LB, GLS, GA, 1PH, TA, 
```

# Module

## Correlation and linear regression

correlation PTS and LB

```{r}
cor(nrl$PTS, nrl$LB)
```
Assess the correlation



Linear Model with total team Line Breaks predicting total team points
```{r}
mod1 <- lm(PTS ~ LB, data = nrl)
summary(mod1)
```
Interpret the intercept. 
Interpret the slope.
Is either the slope or the intercept significant? If so explain the implications.



Correlation between PTS and TA

```{r}
cor(nrl$PTS, nrl$TA)
```
Assess the correlation



Linear model predicting total team points with total team try assists
```{r}
mod2 <- lm(PTS ~ TA, data = nrl)
summary(mod2)
```
Again explain the signficance of intercept and slope.

## Visualizations

```{r}
ggplot(data = nrl,
       aes(x = LB, y = PTS))+
  geom_point()+
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = nrl,
       aes(x = TA, y = PTS))+
  geom_point()+
  geom_smooth(se = FALSE)
```


## Estimation and prediction 

```{r}

```

# CI PI
